### 2018-07-26 ms-thesis-slides.tex
working on slides, writing up abstract for announcement

%!TEX program = xelatex
\documentclass[aspectratio=169]{beamer}

\usepackage{blindtext}

\usetheme{AVThesis}

\title{The Art of Robotics:\\\vspace{0.2em}\huge{Towards A Holistic Approach}}
\subtitle{Master of Science in Robotics\\Thesis Defense}
\author{Alexander Volkov Jr.}
\date{July 31, 2018}

% Change frame margins
\addtobeamertemplate{frametitle}{\vspace*{1cm}}{\vspace*{0cm}}

\setcounter{showSlideNumbers}{1}

\begin{document}
	\setcounter{showProgressBar}{0}
	\setcounter{showSlideNumbers}{0}

	\frame{\titlepage}

	\begin{frame}
		\frametitle{Talk Agenda}
		\begin{enumerate}
			\item Other Things \\ 
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Warm up the crowd with a quick review of some other things I've done.}
			\item The Big Picture \\ 
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Cover the central themes of this talk.}

			\item Bond Graphs, Impedance/Admittance Matching \\
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} An old framework, revived.}

			\item Brushless PMSM Motors \\
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Operating limits, thermal management, liquid cooling.}

			\item Touching Robots \\
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Tactile sensing, whiskers, robot pain.}

			\item Conclusions \\ 
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Review key points and propose some future work.}

			\item Questions \\
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} \emph{``Please clap...''}}
		\end{enumerate}
	\end{frame}

	\setcounter{framenumber}{0}
	\setcounter{showProgressBar}{1}
	\setcounter{showSlideNumbers}{1}

	\section{Other Things}
		\begin{frame}
			\frametitle{Teaching}
			Gelfand Center
		\end{frame}

		\begin{frame}
			\frametitle{Tangential Projects}
			Kruskal count + quadruped gaits
		\end{frame}

		\begin{frame}
			\frametitle{Robotics Institute Meme Facebook Page}
			Started the RI's first and only meme page
			Current stats?
			Some meme examples?
		\end{frame}

	\section{The Big Picture}
		\begin{frame}
			\frametitle{}
			\begin{enumerate}
				\item LaTeX is great!
				\item Beamer is easy to use!
				\item Why not?
			\end{enumerate}
		\end{frame}

	\section{Lorem Ipsum}
		\begin{frame}
			\frametitle{Lorem 1}
			\blindtext
		\end{frame}

		\begin{frame}
			\frametitle{Lorem 2}
			\blindtext
		\end{frame}

		\begin{frame}
			\frametitle{Lorem 3}
			\blindtext
		\end{frame}

	\section{Conclusions}
		\begin{frame}
			\frametitle{Closing Thoughts}
			\begin{itemize}
				\item Woo, Beamer!
			\end{itemize}
		\end{frame}

	\section{Questions?}

	\appendix
	\backupbegin
	  \begin{frame}
	    \frametitle{Backup slide 1}
	    \blindtext
	  \end{frame}
	\backupend

\end{document}