### 2018-07-31 ms-thesis-slides.tex
wow, such progress, much work

%!TEX program = lualatex
\documentclass[aspectratio=169]{beamer}
% add 'handout' to beamer options for printout

\usepackage{multimedia}
\usepackage[backend=biber,style=ieee]{biblatex}

\usetheme{AVThesis}

\title{The Art of Robotics:\\\vspace{0.2em}\huge{Toward a Holistic Approach}}
\subtitle{Master of Science in Robotics\\Thesis Defense}
\author{Alexander Volkov Jr.}
\date{July 31, 2018}

\bibliography{refs/thesis-slides.bib}

% Change frame margins
\addtobeamertemplate{frametitle}{\vspace*{1cm}}{\vspace*{0cm}}

\setcounter{showSlideNumbers}{1}

\begin{document}
	\setcounter{showProgressBar}{0}
	\setcounter{showSlideNumbers}{0}

	\frame{\titlepage}

	\begin{frame}
		\frametitle{Talk Agenda}
		\begin{enumerate}
			\setlength{\itemsep}{4mm}
			\item Background
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} The BFD; how I ended up here}

%			\item Other Things
%				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Warm up the crowd with a quick review of some other things I've done.}

			\item The Big Picture
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Cover the central themes of this talk.}

			\item A Unifying Framework
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} An old framework, revived.}

%			\item Brushless PMSM Motors
%				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Operating limits, thermal management, liquid cooling.}

			\item Touching Robots
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Tactile sensing, contact modeling, whiskers, robot pain.}

			\item Conclusions
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} Review key points and propose some future work.}

			\item Acknowledgments \& Questions
				\textcolor{AVThesisGrey}{\footnotesize\hspace{1em} \emph{``Please clap...''}}
		\end{enumerate}
	\end{frame}

	\setcounter{framenumber}{0}
	\setcounter{showProgressBar}{1}
	\setcounter{showSlideNumbers}{1}

	\section{Background}

		\begin{frame}
			\frametitle{The BFD}
			\begin{itemize}
				\item \textbf{End Game:} Develop a Unified Framework for Understanding Robotics \pause
				\begin{itemize}
					\item \textbf{Thesis Purpose:} Reflect on and organize \emph{my} understanding of robotics \pause
						\begin{itemize}
							\item \textbf{Talk Purpose:} Provide an overview of my thesis dissertation for the committee
						\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Talk Title Etymology I}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=east] (taoe) at ($(current page.east)+(-3mm,-1mm)$) 	{\includegraphics[width=5.5cm]{media/the-art-of-electronics.png}};
			\end{tikzpicture}
			\begin{itemize}
				\item ``The Art of Electronics'', by \emph{Horowitz \& Hill}\pause
				\begin{itemize}
					\item Literally the electrical engineering bible
					\item Incredibly thorough
					\item Perfect balance of practicality and rigor 
				\end{itemize}
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Talk Title Etymology II}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=east] (taoe) at ($(current page.east)+(-10mm,-1mm)$) 	{\includegraphics[width=4cm]{media/mason-manip-book.jpg}};
			\end{tikzpicture}
			\begin{itemize}
				\item ``Mechanics of Robotic Manipulation'', by \emph{Mason}\footnotemark~\autocite{mason2001mechanics} \pause
					\begin{itemize}
						\item ``Manipulation is an art\dots'' (p.1)
					\end{itemize}
			\end{itemize}
			\footnotetext{Matt, I'll take my referral payments by mail.}
		\end{frame}

%		\begin{frame}
%			\frametitle{Talk (Sub)Title Etymology III}
%			\begin{tikzpicture}[overlay, remember picture]
%			\fontsize{18}{15}
%			\node[anchor=center] (subtitle) at (current page.center) {``Toward a Holistic Approach''};
%			\end{tikzpicture}
%		\end{frame}
%		
%		\begin{frame}
%			\addtocounter{framenumber}{-1}
%			\frametitle{Talk (Sub)Title Etymology III}
%			\begin{tikzpicture}[overlay, remember picture]
%				\fontsize{18}{15}
%				\node[anchor=center] (subtitle) at (current page.center) {``\highlight{toward}{Toward} a Holistic Approach''};
%				\fontsize{10}{12}
%				\node[anchor=south west, fill=red!20,rounded corners, text width=3cm, outer sep=3pt, inner sep=3pt] (toward-note) at ($(current page.south west) + (10mm,10mm)$) {When you concede falling short of a goal in academia.};
%				\draw [->, line width=0.3mm] (toward-note) -- (toward);
%			\end{tikzpicture}
%		\end{frame}

		\begin{frame}
%			\addtocounter{framenumber}{-1}
			\frametitle{Talk (Sub)Title Etymology III}
			\begin{tikzpicture}[overlay, remember picture]
				\fontsize{18}{15}
				\only<1>{\node[anchor=center] (subtitle) at (current page.center) {``Toward a Holistic Approach''};}
				\only<2>{\node[anchor=center] (subtitle) at (current page.center) {``\highlight{toward}{Toward} a Holistic Approach''};
				\fontsize{10}{12}
				\node[anchor=south west, fill=red!20,rounded corners, text width=5cm, outer sep=3pt, inner sep=3pt] (toward-note) at ($(current page.south west) + (10mm,10mm)$) {When you concede falling short of a goal in academia. Example: Mason's Annual Review ``Toward Robotic Manipulation'' \autocite{Mason2018}};
				\draw [->, line width=0.3mm] (toward-note) -- (toward);}
				\onslide<3->{\node[anchor=center] (subtitle) at (current page.center) {``\highlight{toward}{Toward} a \highlight{holistic}{Holistic} Approach''};
				\fontsize{10}{12}
				\node[anchor=south west, fill=red!20,rounded corners, text width=5cm, outer sep=3pt, inner sep=3pt] (toward-note) at ($(current page.south west) + (10mm,10mm)$) {When you concede falling short of a goal in academia. Example: Mason's Annual Review ``Toward Robotic Manipulation'' \autocite{Mason2018}};
				\draw [->, line width=0.3mm] (toward-note) -- (toward);
				\node[anchor=south east, fill=red!20,rounded corners, text width=7cm, outer sep=3pt, inner sep=3pt] (holistic-note) at ($(current page.south east) + (-10mm,10mm)$) {From the mechatronics/systems engineering community. Chhabra and Emami provide an excellent summary in \autocite{Chhabra2011}.};
				\draw [->, line width=0.3mm] (holistic-note) -- (holistic);}
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) I}
			\vspace{-0.5em}
			\begin{itemize}
				\item \textbf{Hobby robotics} $\Rightarrow$ self-taught basic electronics, programming, CAD\pause
					\begin{itemize}
						\item ``Keep it cheap''
						\item ``Keep it open-source''
						\item ``Keep it useful (or, at least, artistic)''
					\end{itemize}\pause
				\vspace{0.5em}
				\item \textbf{Cornell Undergrad} $\Rightarrow$ ECE major, ME/CS minors\pause
					\begin{itemize}
						\item \textbf{ECE:}
							\begin{itemize}
								\item ``Everything is an impedance (or an admittance)''
								\item ``What's the system bandwidth?''
							\end{itemize}
						\item \textbf{ME:}\pause
							\begin{itemize}
								\item ``At the end of the day, everything is mechanical''
								\item ``Everything is a model, and every model is wrong''
							\end{itemize}\pause
						\item \textbf{CS:} 
						\begin{itemize}
							\item ``90\% of solving a problem is finding the right representation''
							\item ``Everything breaks at the interfaces''
						\end{itemize}
					\end{itemize}
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) II}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=center] (ri-logo) at (current page.center) {\includegraphics[width=30mm]{assets/ri-logo.eps}};
				\node[minimum width=2cm, below=5mm of ri-logo] {``Time to become a robotics master...''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) III}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{10}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {``Legged locomotion is cool...''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) IV}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{10}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {\color{gray}``Legged locomotion is cool...''};
				\node[anchor=north east] (hartmut) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/hartmut.jpg}};
				\node[anchor=north east] (travers) at ($(current page.north east)+(-28mm,-15mm)$) {\includegraphics[width=2cm]{media/travers.jpg}};
				\node[anchor=north east] (koushil) at ($(current page.north east)+(-51mm,-15mm)$) {\includegraphics[width=2cm]{media/koushil.jpg}};
				\node[anchor=east, minimum width=2cm, left=3mm of koushil] {``Great! Here's absolutely no funding.''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) V}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{10}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {``Umm...''};
				\node[anchor=north east] (hartmut) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/hartmut.jpg}};
				\node[anchor=north east] (travers) at ($(current page.north east)+(-28mm,-15mm)$) {\includegraphics[width=2cm]{media/travers.jpg}};
				\node[anchor=north east] (koushil) at ($(current page.north east)+(-51mm,-15mm)$) {\includegraphics[width=2cm]{media/koushil.jpg}};
				\node[anchor=east, minimum width=2cm, left=3mm of koushil] {\color{gray}``Great! Here's absolutely no funding.''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) VI}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{10}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {\color{gray}``Umm...''};
				\node[anchor=north east] (ross) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/ross.jpg}};
				\node[anchor=east, minimum width=2cm, left=3mm of ross] {``Wait! Go talk to Matt Mason!''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) VII}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{10}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {``(skeptical) Okay...''};
				\node[anchor=north east] (ross) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/ross.jpg}};
				\node[anchor=east, minimum width=2cm, left=3mm of ross] {\color{gray}``Wait! Go talk to Matt Mason!''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) VIII}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{15}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {\color{gray}``(skeptical) Okay...''};
				\node[anchor=north east] (matt) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/mason.jpg}};
				\node[anchor=east, text width=5cm, left=3mm of matt, align=right] {``Manipulation is awesome! \\And I have money! \\And I don't micromanage!''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) IX}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{15}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {``(excited) Works for me!''};
				\node[anchor=north east] (matt) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/mason.jpg}};
				\node[anchor=east, text width=5cm, left=3mm of matt, align=right] {\color{gray}``Manipulation is awesome! \\And I have money! \\And I don't micromanage!''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) X}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{15}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[minimum width=2cm, right=3mm of me] {\color{gray}``(excited) Works for me!''};
				\node[anchor=north east] (matt) at ($(current page.north east)+(-5mm,-15mm)$) {\includegraphics[width=2cm]{media/mason.jpg}};
				\node[anchor=east, minimum width=2cm, left=3mm of matt, align=right] {``Fantastic, go forth and prosper!''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) XI}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=south east] (ri-logo) at ($(current page.south east)+(-5mm,10mm)$) {\includegraphics[width=16mm]{assets/ri-logo.eps}};
				\fontsize{12}{15}
				\node[anchor=south west] (me) at ($(current page.south west)+(5mm,10mm)$) {\includegraphics[width=2cm]{media/me.jpg}};
				\node[text width=8cm, right=3mm of me, align=left] {``I've been meaning to read Hogan's famous \emph{Impedance Control}~\autocite{hogan1985impedance} paper, I guess I'll start there...''};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
			\frametitle{How I Ended Up Here (Some Context) XII}
			\begin{tikzpicture}[overlay, remember picture]
				\fontsize{18}{15}
				\node[anchor=center, minimum width=2cm] at (current page.center) {\dots~Two Years Later \dots};
			\end{tikzpicture}
		\end{frame}

		\begin{frame}
		\frametitle{Me, circa July 2018}
			\begin{tikzpicture}[overlay, remember picture]
				\node[anchor=center] (paranoid) at (current page.center) {\includegraphics[width=8cm]{media/charlie-paranoid}};
				\node[anchor=center, below=0mm of paranoid] {``So it's all about causality\dots~and feedback!''};
			\end{tikzpicture}
		\end{frame}

	\section{The Big Picture}
		\begin{frame}
			\frametitle{Back to Basics}
			\begin{itemize}[<+->]
				\item Want a concise \emph{theory of robotics}\dots
				\item Interested in the \emph{physics} common to robotics problems
				\item Not worried about SLAM, POMDPs, etc\dots~there are bigger fish to fry (we're skimming over the basics!)
					\begin{itemize}
						\item i.e. Moravec's paradox \autocite{moravec1988mind}
					\end{itemize}
				\item Inherently a \emph{breadth-first} approach, since we're looking for a unifying framework
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Problem 1: What is a Robot?}
			\begin{itemize}
				\item I know it's a cliché to bring this up\dots \pause but I must! \pause
				\begin{itemize}
					\item Does the ambiguity really matter? \pause
					\item Maybe not, but some unifying \emph{theme} would be useful!
				\end{itemize}
			\end{itemize}
		\end{frame}

		\begin{frame}

		\end{frame}

		\begin{frame}
			\frametitle{Problem 2: Locomotion and Manipulation, Segregated}
			\begin{itemize}
				\item Locomotion and manipulation are studied separately in robotics\pause
					\begin{itemize}
						\item \dots~and biomechanics, for that matter
						\item Seems quite natural at first, we all talk about the two as separate specializations in robotics
					\end{itemize}\pause
				\item Eventually, the notion of \emph{``duality''} comes up\dots
				\begin{itemize}\pause
					\item Locomotion and manipulation sometimes overlap
						\begin{itemize}
							\item Pai et al.'s \emph{Platonic Beasts} \autocite{pai1994platonic}
							\item Mason et al.'s \emph{Mobipulator} \autocite{MasonMobileManipulator1999} \autocite{MasonDesktopMobileManipulators2000}
							\item Also, literally everywhere in biology
						\end{itemize}
					\item Perhaps it just comes down to a change of reference?
						\begin{itemize}
							\item What pushes off of what?
							\item Locomotion is ``self-manipulation'', e.g. Aaron Johnson's PhD thesis \autocite{johnson2014self} and related works \autocite{johnson2012standing}\autocite{johnson2016hybrid}
						\end{itemize}
				\end{itemize}
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Problem 2: Locomotion and Manipulation, United}
			\begin{itemize}
				\item Taking the ``self-manipulation'' perspective
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Embodiment \& Hogan's Physical-Equivalence Principle}
			\begin{itemize}[<+->]
				\item Embodiment
					\begin{itemize}
						\item 
						\item Presently resides in the cognitive science community (as in ``embodied \emph{cognition}'')
					\end{itemize}
				\item 
			\end{itemize}
		\end{frame}

	\section{A Unifying Framework}
		\begin{frame}
			\frametitle{The Port-Hamiltonian Framework}
			\begin{itemize}
				\item Port-based Analysis
				\item Hamiltonian Dynamics
					\begin{itemize}
						\item 
					\end{itemize}
				\item
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{Bond Graphs}

		\end{frame}

		\begin{frame}
			\frametitle{Hamiltonian Dynamics}

		\end{frame}

		\begin{frame}
			\frametitle{Dissipative Elements}
		\end{frame}

		\begin{frame}
			\frametitle{Putting it All Together}

		\end{frame}

	\section{Touching Robots}

		\begin{frame}
			\frametitle{A Note on Stiffness and Collisions}
			\begin{itemize}
				\item \textbf{\color{red}STOP COLLIDING STIFF OBJECTS AND EXPECTING RELIABLE RESULTS}
			\end{itemize}
		\end{frame}

		\begin{frame}
			\frametitle{The Most Important Sensory Modality}
		\end{frame}

		\begin{frame}
			\frametitle{Contact Models}
		\end{frame}

		\begin{frame}
			\frametitle{Terminator 2 Got It Right}
			\begin{center}
			\movie[height = 0.6\textwidth, width = 0.8\textwidth,showcontrols]{}{media/terminator2-pain.mp4}
			\end{center}
		\end{frame}

	\section{Closing Thoughts}
		\begin{frame}
			\frametitle{Closing Thoughts}
			\begin{itemize}
				\item Woo, Beamer!
			\end{itemize}
		\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\setcounter{showProgressBar}{0}
	\setcounter{showSlideNumbers}{0}

	\appendix
	\backupbegin

	\section{Acknowledgments}

	\begin{frame}
	\frametitle{Cheers}
		Committee\\Lab\\The other profs\\Jean\\Cameron\\Cats\\Parents for debt free education
	\end{frame}

	\section{Questions?}

	\nocite{*}
	\begin{frame}[allowframebreaks]
		\frametitle{Bibliography}
		\printbibliography[heading=none]
	\end{frame}

		\section{Other Things}
	\begin{frame}
	\frametitle{Teaching with CMU's Gelfand Outreach Center}
	\begin{tikzpicture}[overlay, remember picture]
	\node[anchor=south east] (gelfand-logo) at ($(current page.south east) + (-8mm, 8mm)$) {\includegraphics[width=4cm]{media/gelfand-logo.jpg}};
	\end{tikzpicture}
	\vspace{9em}
	\begin{itemize}
		\item Introduction to Robotics with the Finch Platform
		\begin{itemize}
			\item Each session is 5 days, 3 hrs/day
			\item 4\textsuperscript{th} and 5\textsuperscript{th} graders
			\item 1 session Summer 2017 + 2 sessions Summer 2018
		\end{itemize}
	\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Teaching with CMU's Gelfand Outreach Center}
\begin{tikzpicture}[overlay, remember picture]
\node[anchor=south east] (gelfand-logo) at ($(current page.south east) + (-8mm, 8mm)$) {\includegraphics[width=4cm]{media/gelfand-logo.jpg}};
\end{tikzpicture}
\vspace{9em}
\begin{itemize}
\item Saturday Series LEGO WeDo Robotics
\begin{itemize}
	\item 3 hour course
	\item 2\textsuperscript{nd} and 3\textsuperscript{rd} graders
	\item 1 session Spring 2018
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Robotics Institute Meme Facebook Page}
\begin{tikzpicture}[overlay, remember picture]
\node[anchor=center] (ri-meme-screenshot) at (current page.center) {\includegraphics[width=\textwidth]{media/ri-memes.jpg}};
\end{tikzpicture}
\end{frame}

\begin{frame}
\frametitle{Robotics Institute Meme Facebook Page}
\vspace{9em}
\begin{itemize}
\item Some statistics:
\begin{itemize}
\item The RI's first and only meme page
\item Formed in March 2018
\item 170 members
\item 30 memes, 23 original contributions!
\end{itemize}
\end{itemize}
\begin{tikzpicture}[overlay, remember picture]
\node[anchor=south east] (ri-meme-3) at ($(current page.south east) + (2mm,5mm)$) {\includegraphics[width=7.5cm]{media/ri-meme-manifold-optimization.jpg}};
\end{tikzpicture}
\begin{tikzpicture}[overlay, remember picture]
\node[anchor=north east] (ri-meme-1) at ($(current page.north east) + (-45mm,-12mm)$) {\includegraphics[width=3.5cm]{media/ri-meme-thesis-anxiety.jpg}};
\end{tikzpicture}
\begin{tikzpicture}[overlay, remember picture]
\node[anchor=north west] (ri-meme-2) at ($(current page.north west) + (10mm,-12mm)$) {\includegraphics[width=5cm]{media/ri-meme-ros-sleep.jpg}};
\end{tikzpicture}
\end{frame}

	\backupend

\end{document}